---
title: "Project 1" 
author: "Julian N"
date: "2024-02-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = "")
library(caret)
library(randomForest)
library(readr)
library(dplyr)
library(plotly)
library(palmerpenguins)
```

# Introduction to the Data

The goal of this project is to create a classification model using the Palmer Penguins data Set,

and to visualize the accuracy of different methods.

```{r}

```

## Processing data

```{r split, message=FALSE, warning=FALSE, include=FALSE, paged.print=TRUE}

set.seed(5339)
levels(penguins$species)
 penguins <- na.omit(penguins)
     penguins<- subset(penguins, select = -sex)
      penguins <- subset(penguins, select = -island)
      


```

```{r}
head(penguins)
```

After removing all NA's, the "sex" column, and "island" column, the data set consists of a list of penguin species, and identifying factors like like flipper length and body mass.

```{r}
trainingSet <-createDataPartition(penguins$species, p=0.80,list = FALSE)


training <-penguins[trainingSet,]

    tester <- penguins[-trainingSet,]#remaining

```

Using the *caret* library I then subset the data into a training and a testing set, with 20% of the data reserved for validating the results

## Making a model

```{r algorithm }
penguinC<- trainControl(method='cv', number =10)

metric = 'Accuracy'
#linear discriminant analysis
penguinTrainingLDA <- train(
                        species ~ .,
                        data= penguins,
                        metric=metric,
                        trconrol=penguinC,
                        method = 'lda')

penguinPredictions<- predict(penguinTrainingLDA,tester)



```

after training the model using the Linear Discriminant Analysis method we utilize the predict function to cross reference with our test dataset.

## Viewing the Model

```{r}
head(penguinPredictions)
```

The predictions are not very very view-able, so we can pass it into a confusion matrix to see how the model did.

```{r matrix, message=TRUE}


#levels(penguinPredictions)
#levels(as.factor(tester$species))
##as.facor changes levels to same as predictions


confusionMatrix(penguinPredictions,as.factor(tester$species))
```

😀The model seems to have gotten all of the test predictions correct

## Testing other Methods

```{r}

set.seed(5339)



penguinTrainingRF <- train(
                        species ~ .,
                        data= penguins,
                        metric=metric,
                        trconrol=penguinC,
                        method = "rf")

predictionRF <- predict(penguinTrainingRF,tester)

confusionMatrix(predictionRF,as.factor(tester$species))



```

We preform the same steps using the Random Forest training method, however on the surface there seems to be no difference in accuracy.

## Resamples

```{r plottingrs, echo=FALSE, warning=FALSE}
ldaPlot<-plot_ly(penguinTrainingLDA$resample,
        x = ~Resample, 
        y = ~Accuracy,
        mode = "markers",
        type = "scatter"
        )


ldaPlot<-ldaPlot %>% layout(
  title = 'LDA Resamples',
  xaxis = list(title = 'sample', showticklabels = FALSE),  
  yaxis = list(title = 'accuracy')
)
ldaPlot
```

Viewing the re-samples individually for the models paints a more accurate picture of the models and allows for further understanding of the effectiveness of the models

## Comparison of Re-samps

```{r ldrfplotted, echo=FALSE, warning=FALSE}

combinedData <- bind_rows(
   mutate(penguinTrainingLDA$resample, model = "LDA"),
  mutate(penguinTrainingRF$resample, model = "Random Forest")
 
)

combinedPlot <- plot_ly(combinedData, x = ~Resample, y = ~Accuracy, color = ~model, type = "scatter", mode = "markers")
  
combinedPlot<-combinedPlot%>%layout(
  title = "LDA and RF comparison",
 xaxis = list(title = 'sample', showticklabels = FALSE)
)

combinedPlot
```

When re-samples from both RF and LDA are plotted we see that LDA has significantly more samples with 100% accuracy.

### Averages

```{r echo=FALSE}
penguinResamps <- resamples(list(
  lda = penguinTrainingLDA,
  rf =penguinTrainingRF ))
summary(penguinResamps)
```

We can see that in the re-samples that the lda method had slightly higher accuracy with a 98% mean whereas rf had a 97% mean

```{r echo=FALSE}

ggplot(data = penguinResamps)+
  theme_dark()
```

We can better visualize the difference in accuracy when we plot this data and can conclude that LDA is more accurate than RF in this application, and if only marginally.
